---
author: "Gilverto De Los Santos Rios"
output: html_document
source: https://www.kaggle.com/datasets/fivethirtyeight/uber-pickups-in-new-york-city
---

```{r}
library(lubridate)
library(ggplot2)
```
# Load Data
```{r}
april = read.csv(file = "~/R/projects/Uber/data/uber-raw-data-apr14.csv", header = TRUE, sep = ",")
aug = read.csv(file = "~/R/projects/Uber/data/uber-raw-data-aug14.csv", header = TRUE, sep = ",")
jul = read.csv(file = "~/R/projects/Uber/data/uber-raw-data-jul14.csv", header = TRUE, sep = ",")
jun = read.csv(file = "~/R/projects/Uber/data/uber-raw-data-jun14.csv", header = TRUE, sep = ",")
may = read.csv(file = "~/R/projects/Uber/data/uber-raw-data-may14.csv", header = TRUE, sep = ",")
sep = read.csv(file = "~/R/projects/Uber/data/uber-raw-data-sep14.csv", header = TRUE, sep = ",")
```


# Trip Information for April
```{r}
april$Date.Time = as.POSIXct(april$Date.Time, format = "%m/%d/%Y %H:%M:%S")
april$day = day(april$Date.Time)
april$hour = hour(april$Date.Time)
aprilDays = as.data.frame(table(april$day))
aprilHours = as.data.frame(table(april$hour))
```

# Plot Trips for April
```{r}
ggplot(aprilDays, aes(x = Var1, y = Freq)) + 
  geom_bar(stat = "identity", fill = "skyblue") + 
  labs(title = "Trips per Day in April", x = "Day of the Month", y = "Trips")

ggplot(aprilHours, aes(x = Var1, y = Freq)) + 
  geom_bar(stat = "identity", fill = "skyblue") + 
  labs(title = "Trips per Hour in April", x = "Hour", y = "Trips")
```

# Trip Information for August
```{r}
aug$Date.Time = as.POSIXct(aug$Date.Time, format = "%m/%d/%Y %H:%M:%S")
aug$day = day(aug$Date.Time)
aug$hour = hour(aug$Date.Time)
augDays = as.data.frame(table(aug$day))
augHours = as.data.frame(table(aug$hour))
```

# Plot Trips for August
```{r}
ggplot(augDays, aes(x = Var1, y = Freq)) + 
  geom_bar(stat = "identity", fill = "skyblue") + 
  labs(title = "Trips per Day in August", x = "Day of the Month", y = "Trips")

ggplot(augHours, aes(x = Var1, y = Freq)) + 
  geom_bar(stat = "identity", fill = "skyblue") + 
  labs(title = "Trips per Hour in August", x = "Hour", y = "Trips")
```

# Trips Information for July
```{r}
jul$Date.Time <- as.POSIXct(jul$Date.Time, format = "%m/%d/%Y %H:%M:%S")
jul$day <- day(jul$Date.Time)
jul$hour <- hour(jul$Date.Time)
julDays <- as.data.frame(table(jul$day))
julHours <- as.data.frame(table(jul$hour))
```

# Plot Trips for July
```{r}
ggplot(julDays, aes(x = Var1, y = Freq)) + 
  geom_bar(stat = "identity", fill = "skyblue") + 
  labs(title = "Trips per Day in July", x = "Day of the Month", y = "Trips")

ggplot(julHours, aes(x = Var1, y = Freq)) + 
  geom_bar(stat = "identity", fill = "skyblue") + 
  labs(title = "Trips per Hour in July", x = "Hour", y = "Trips")
```

# Trip Information for June
```{r}
jun$Date.Time <- as.POSIXct(jun$Date.Time, format = "%m/%d/%Y %H:%M:%S")
jun$day <- day(jun$Date.Time)
jun$hour <- hour(jun$Date.Time)
junDays <- as.data.frame(table(jun$day))
junHours <- as.data.frame(table(jun$hour))
```

# Plot Trips for June
```{r}
ggplot(junDays, aes(x = Var1, y = Freq)) + 
  geom_bar(stat = "identity", fill = "skyblue") + 
  labs(title = "Trips per Day in June", x = "Day of the Month", y = "Trips")

ggplot(junHours, aes(x = Var1, y = Freq)) + 
  geom_bar(stat = "identity", fill = "skyblue") + 
  labs(title = "Trips per Hour in June", x = "Hour", y = "Trips")
```

# Trip Information for May
```{r}
may$Date.Time <- as.POSIXct(may$Date.Time, format = "%m/%d/%Y %H:%M:%S")
may$day <- day(may$Date.Time)
may$hour <- hour(may$Date.Time)
mayDays <- as.data.frame(table(may$day))
mayHours <- as.data.frame(table(may$hour))
```

# Plot Trips for May
```{r}
ggplot(mayDays, aes(x = Var1, y = Freq)) + 
  geom_bar(stat = "identity", fill = "skyblue") + 
  labs(title = "Trips per Day in May", x = "Day of the Month", y = "Trips")

ggplot(mayHours, aes(x = Var1, y = Freq)) + 
  geom_bar(stat = "identity", fill = "skyblue") + 
  labs(title = "Trips per Hour in May", x = "Hour", y = "Trips")
```

# Trip Information for September
```{r}
sep$Date.Time <- as.POSIXct(sep$Date.Time, format = "%m/%d/%Y %H:%M:%S")
sep$day <- day(sep$Date.Time)
sep$hour <- hour(sep$Date.Time)
sepDays <- as.data.frame(table(sep$day))
sepHours <- as.data.frame(table(sep$hour))
```

# Plot Trips for September
```{r}
ggplot(sepDays, aes(x = Var1, y = Freq)) + 
  geom_bar(stat = "identity", fill = "skyblue") + 
  labs(title = "Trips per Day in September", x = "Day of the Month", y = "Trips")

ggplot(sepHours, aes(x = Var1, y = Freq)) + 
  geom_bar(stat = "identity", fill = "skyblue") + 
  labs(title = "Trips per Hour in September", x = "Hour", y = "Trips")
```

# Trip Information for all Months
```{r}
combinedData = rbind(april, aug, jul, jun, may, sep)
combinedData$month = month(combinedData$Date.Time)
combinedDataMonths = as.data.frame(table(combinedData$month))
combinedDataHours = as.data.frame(table(combinedData$hour))
```

# Plots for Combined Months
```{r}
ggplot(combinedDataMonths, aes(x = Var1, y = Freq)) + 
  geom_bar(stat = "identity", fill = "skyblue") + 
  labs(title = "Trips per Month", x = "Month", y = "Trips")

ggplot(combinedDataHours, aes(x = Var1, y = Freq)) + 
  geom_bar(stat = "identity", fill = "skyblue") + 
  labs(title = "Trips by Hour for Six Months", x = "Time", y = "Trips")
```




